[Artifact Presentation] Damegender: \\ Writing and Comparing Gender Detection Tools

### 2020-05-15 paper2sattose.tex
Trabajando en la parte de ML

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{twocolceurws}
\usepackage{xcolor}
\usepackage{url}

@@

\author{
David Arroyo Menéndez, Jesus González-Barahona, Gregorio Robles \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc, grex@gsyc\}.urjc.es
}

@@

In Section 2 we explain \texttt{damegender} as solution to the
problem of gender guessing. Section 3 introduces the Open Datasets used as the \emph{golden truth}.
Section 4 presents a feature comparison with other gender guession tools and services.
Section 5 reports on accuracies and offers a confusion matrix using a scientific dataset.
Section 6 is about how we use Machine Learning in \texttt{damegender}.
Section 7 discusses limitations and futher research, and concludes the paper.

@@

\section{Reproducing accuracies and confusion matrix}

@@

%\begin{table*}[]{@{}lllllll@{}}
\begin{table}[t]
\footnotesize
\begin{tabular}[]{lcccc}
  \hline
  API & Acc & Prec & F1 & Recall\tabularnewline
\hline
Genderapi & 0.969 & 0.972 & 0.964 & 1.0\tabularnewline
Genderize & 0.927 & 0.976 & 0.966 & 1.0\tabularnewline
Damegender (SVC)\footnote{SVC is the acronym of Support Vector Classification, the Machine Learning algorithm that Damegender was using with this results} & 0.879 & 0.972 & 0.972 & 1.0\tabularnewline
Namsor & 0.867 & 0.973 & 0.924 & 1.0\tabularnewline
Nameapi & 0.830 & 0.974 & 0.905 & 1.0\tabularnewline
Gender Guesser & 0.774 & 0.985 & 0.872 & 1.0\tabularnewline
\hline
\end{tabular}
\caption{Different accuracies measures}
\label{table:DifferentAccuraciesMeasures}
\end{table}

We have performed a comparison using a confusion matrix for the software/tools (see Table~\ref{table:ConfusionMatrixTables}).
Compared to the results obtained in~\cite{10.7717/peerj-cs.156}, we can see that they are very similar.
The most important tools (\texttt{Namsor}, \texttt{Genderapi} and \texttt{Genderize}) are improving the accuracies with respect the previous comparison.
In particular, \texttt{Genderapi} has similar results, but it improves the results for \emph{undefined}.
In \texttt{Genderguesser} we obtain different results, which is to some extent not expected, because the software has not modified for several years.
For \texttt{Genderize}, we obtain the same results. 
\texttt{Nameapi}'s results is changing from male to female with more errors. 
In \texttt{Namsor}, the results are similar. 
\texttt{damegender} is not guessing undefined because we predict with machine learning (SVC) if the string is not in the database.

@@

@@

\begin{table*}
\footnotesize
\center
\begin{tabular}[]{lccccccc}
  \hline
Dataset & contains a & last is a & last is o & last is consonant & last is vocal & 1st is consonant & 1st is vocal  \tabularnewline
  \hline
  %\endhead
 Uruguay (F) \footnote{F is for females and M is for males} &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
 Uruguay (M) \footnotemark[3] &    0.643 &         0.249 &         0.062 &                 0.766 &             0.234 &                  0.771 &              0.228  \tabularnewline
 Spain (F) \footnotemark[3]  &    0.922 &         0.588 &          0.03 &                 0.271 &             0.728 &                  0.772 &              0.228  \tabularnewline
 Spain (M) \footnotemark[3]  &    0.818 &          0.03 &         0.268 &                 0.569 &              0.43 &                  0.763 &              0.236  \tabularnewline
 UK (F) \footnotemark[3]     &    0.825 &         0.374 &         0.013 &                 0.322 &             0.674 &                  0.765 &              0.235  \tabularnewline
 UK (M) \footnotemark[3]     &    0.716 &         0.036 &         0.039 &                  0.78 &             0.218 &                  0.799 &                0.2  \tabularnewline
 USA (F) \footnotemark[3]    &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
 USA (M) \footnotemark[3]    &    0.643 &          0.02 &         0.061 &                 0.765 &             0.234 &                   0.84 &              0.159  \tabularnewline
 Canada (F) \footnotemark[3] &    0.659 &         0.189 &         0.005 &                 0.591 &             0.408 &                  0.838 &              0.16  \tabularnewline
 Canada (M) \footnotemark[3] &    0.752 &          0.22 &         0.025 &                  0.54 &             0.456 &                  0.818 &              0.181  \tabularnewline
\hline
\end{tabular}
\caption{Informative Features in Different Countries. F stands for females, and M for males.}
\vspace{0.3cm}
\label{table:InfoFeatures}
\end{table*}

@@

The relevant metrics with the different algorithms are shown in Table~\ref{table:MLAccuracies}.

\hline
\end{tabular}
\caption{Machine Learning Algorithms and accuracies measures}
\label{table:MLAccuracies}
\end{table}

The results in Table~\ref{table:MLAccuracies} show that using algorithms as Support Vector Machines or Random Forest against a scientific dataset created by independent researchers where it is possible to reach results similar to another commercial solutions about gender detection tools.
Our classifier is binary (only male and female).

We were doing this experiment with NLTK and INE datasets with accuracies reaching accuracies until 0.745.
So it makes sense expect better results in random datasets augmenting languages and countries.
However, our solution is not providing Arabic or Chinese alphabets, yet.

Based on this experiment, we think that it is possible to infer that \texttt{damegender} could provide a good solution for nicknames, diminutives, or similar.

@@

Thanks to:

Lucía Santamaría and Helena Mihaljevic by the previous work.

Daniel Izquierdo and Laura Arjona for starts this research field in the URJC.

Jesus González Boticario and the people who is working with him in the UNED by the motivation towards the machine learning.

All people working with Jesus González Barahona and Gregorio Robles.

>>>>

\documentclass[a4paper]{article}
\usepackage{graphicx}
\usepackage{twocolceurws}
\usepackage{xcolor}
\usepackage{url}

@@

\author{
David Arroyo Menendez, Jesus Gonzalez-Barahona, Gregorio Robles \\ Grupo de Sistemas y Comunicaciones (GSyC) \\ Universidad Rey Juan Carlos, Madrid, Spain \\ \{d.arroyome@alumnos, jgb@gsyc, grex@gsyc\}.urjc.es
}

@@

In Section 2 we explain \texttt{damegender} as solution to the
problem of gender guessing. Section 3 introduces the Open Datasets used as the \emph{golden truth}.
Section 4 presents a feature comparison with other gender guession tools and services.
Section 5 reports on values of accuracy and offers a confusion matrix using a scientific dataset.
Section 6 is about how we use Machine Learning in \texttt{damegender}.
Section 7 discusses limitations and futher research, and concludes the paper.

@@

\section{Reproducing values of accuracy and confusion matrix}

@@

%\begin{table*}[]{@{}lllllll@{}}
\begin{table}[t]
\footnotesize
\begin{tabular}[]{lcccc}
  \hline
  API & Acc & Prec & F1 & Recall\tabularnewline
\hline
Genderapi & 0.969 & 0.972 & 0.964 & 1.0\tabularnewline
Genderize & 0.927 & 0.976 & 0.966 & 1.0\tabularnewline
Damegender (SVC)\footnote{SVC is the acronym of Support Vector Classification, the Machine Learning algorithm that Damegender was using with this results} & 0.879 & 0.972 & 0.972 & 1.0\tabularnewline
Namsor & 0.867 & 0.973 & 0.924 & 1.0\tabularnewline
Nameapi & 0.830 & 0.974 & 0.905 & 1.0\tabularnewline
Gender Guesser & 0.774 & 0.985 & 0.872 & 1.0\tabularnewline
\hline
\end{tabular}
\caption{Comparison of measures of the quality of the results for the tools under study.}
\label{table:DifferentAccuraciesMeasures}
\end{table}

We have performed a comparison using a confusion matrix for the software/tools (see Table~\ref{table:ConfusionMatrixTables}).
Compared to the results obtained in~\cite{10.7717/peerj-cs.156}, we can see that they are very similar.
The most important tools (\texttt{Namsor}, \texttt{Genderapi} and \texttt{Genderize}) are improving the values of accuracy with respect the previous comparison.
In particular, \texttt{Genderapi} has similar results, but it improves the results for \emph{undefined}.
In \texttt{Genderguesser} we obtain different results, which is to some extent not expected, because the software has not modified for several years.
For \texttt{Genderize}, we obtain the same results. 
\texttt{Nameapi}'s results is changing from male to female with more errors. 
In \texttt{Namsor}, the results are similar. 
\texttt{damegender} is not guessing undefined because we predict with machine learning (SVC) if the string is not in the database.

@@

\subsection{Comparing ML algorithms}

\hline
\end{tabular}
\caption{Machine Learning Algorithms and accuracy measures}
\label{table:MLAccuracies}
\end{table}

Table~\ref{table:MLAccuracies} shows the accuracy measures for some Machine Learning algorithms used for in our guessing. 
The best results are given for Support Vector Machines and Random Forest -- with those algorithms \texttt{damegender} achieves values that are close to more mature, proprietary solutions.
Our classifier is binary (only male and female).

\subsection{Experimenting with some features}

@@

\begin{table*}
\footnotesize
\center
\begin{tabular}[]{lccccccc}
  \hline
Dataset & contains a & last is a & last is o & last is consonant & last is vocal & 1st is consonant & 1st is vocal  \tabularnewline
  \hline
  %\endhead
 Uruguay (F) &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
 Uruguay (M) &    0.643 &         0.249 &         0.062 &                 0.766 &             0.234 &                  0.771 &              0.228  \tabularnewline
 Spain (F)  &    0.922 &         0.588 &          0.030 &                 0.271 &             0.728 &                  0.772 &              0.228  \tabularnewline
 Spain (M)  &    0.818 &         0.030 &         0.268 &                 0.569 &              0.430 &                  0.763 &              0.236  \tabularnewline
 UK (F)     &    0.825 &         0.374 &         0.013 &                 0.322 &             0.674 &                  0.765 &              0.235  \tabularnewline
 UK (M)     &    0.716 &         0.036 &         0.039 &                 0.780 &             0.218 &                  0.799 &                0.200  \tabularnewline
 USA (F)    &    0.816 &         0.456 &         0.007 &                 0.287 &             0.712 &                  0.823 &              0.177  \tabularnewline
 USA (M)    &    0.643 &          0.020 &         0.061 &                 0.765 &             0.234 &                   0.840 &              0.159  \tabularnewline
\hline
\end{tabular}
\caption{Informative Features in Different Countries. F stands for females, and M for males.}
\vspace{0.3cm}
\label{table:InfoFeatures}
\end{table*}

@@

We have done this experiment with the NLTK and INE datasets, with the values of accuracy reaching up to 0.745.
So it makes sense to expect better results in random datasets if we add new languages and countries.
However, our solution is not providing Arabic or Chinese alphabets, yet.
The results of this experiment could be used to provide a good solution for nicknames, diminutives, or similar.

@@

Daniel Izquierdo and Laura Arjona for starting this research field at the URJC.

To Jes\'us Gonz\'alez Boticario and the people who work with him at UNED, because of the motivation for the machine learning part.
